---
title: "Penuel The Black Pen - YT Channel descriptive analysis"
date:  "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
  word_document: default
---

```{r message=FALSE,warning=FALSE}
library(httr)
library(dplyr)
library(jsonlite)
```


Outline: 

1. Pulling relevant data from YT API

2. Testing the "affiliation effect"

3. Exploratory Analysis

4. Top performing videos

5. Closing remarks

# Pulling relevant data from YT API

```{r, warning=FALSE,message=FALSE}

## Sample code for extracting YouTube Data in R Penuel's channel using username

# input your youtube API key

key = "AIzaSyB7gCrBBUVi775tGiec9LcKMs60Aa7JqH4"

# setting up variables, that'll be frequently used
channel_id <- "UCtdTwmNiROeEqmv_SGk513g"  # God Penuel Channel ID
base <- "https://www.googleapis.com/youtube/v3/"

# Construct the API call
# Construct the API call
api_params <- 
  paste(paste0("key=", key), 
        paste0("id=", channel_id), 
        "part=snippet,contentDetails,statistics",
        sep = "&")
api_call <- paste0(base, "channels", "?", api_params)
api_result <- GET(api_call)
json_result <- content(api_result, "text", encoding="UTF-8")

# Process the raw data into a data frame
channel.json = fromJSON(json_result, flatten = T)
channel.df = as.data.frame(channel.json)

playlist_id <- channel.df$items.contentDetails.relatedPlaylists.uploads

```

# Processing the videos from the channel

```{r warning=FALSE,message=FALSE}

# temporary variables
nextPageToken = ""
upload.df = NULL
pageInfo = NULL

# Loop through the playlist while there is still a next page
while (!is.null(nextPageToken)) {
  # Construct the API call
  api_params <- 
    paste(paste0("key=", key), 
          paste0("playlistId=", playlist_id), 
          "part=snippet,contentDetails,status",
          "maxResults=50",
          sep = "&")
  
  # Add the page token for page 2 onwards
  if(nextPageToken != "") {
    api_params <- paste0(api_params,
                         "&pageToken=",nextPageToken)
  }
  
  api_call <- paste0(base, "playlistItems", "?", api_params)
  api_result <- GET(api_call)
  json_result <- content(api_result, "text", encoding="UTF-8")
  upload.json <- fromJSON(json_result, flatten = T)
  
  nextPageToken <- upload.json$nextPageToken
  pageInfo <- upload.json$pageInfo
  
  curr.df <- as.data.frame(upload.json$items)
  if (is.null(upload.df)) {
    upload.df <- curr.df
  } else {
    upload.df <- bind_rows(upload.df, curr.df)
  }
}

video.df<- NULL
# Loop through all uploaded videos
for (i in 1:nrow(upload.df)) {
  # Construct the API call
  video_id <- upload.df$contentDetails.videoId[i]
  api_params <- 
    paste(paste0("key=", key), 
          paste0("id=", video_id), 
          "part=id,statistics,contentDetails",
          sep = "&")
  
  api_call <- paste0(base, "videos", "?", api_params)
  api_result <- GET(api_call)
  json_result <- content(api_result, "text", encoding="UTF-8")
  video.json <- fromJSON(json_result, flatten = T)
  
  curr.df <- as.data.frame(video.json$items)
  
  if (is.null(video.df)) {
    video.df <- curr.df
  } else {
    video.df <- bind_rows(video.df, curr.df)
  }
}  

# Combine all video data frames
video.df$contentDetails.videoId <- video.df$id
video_final.df <- merge(x = upload.df, 
                        y = video.df,
                        by = "contentDetails.videoId")

write.csv(x = channel.df,
          row.names = F,
          file = "God_Penuel_Channel_updated.csv")
write.csv(x = video_final.df,
          row.names = F,
          file = "God_Penuel_Uploads_updated.csv")

```


# Reading in the data
```{r warning=FALSE,message=FALSE}
##------------------------------Reading-In-The-Data---------------------------##

# Loading in the relevant packages
library(readxl)
library(tidyverse)
library(lubridate)

# generate a full path to a file
filename = "God_Penuel_Uploads_updated.csv"

full_path = file.path(getwd(), filename)

gp_data = read.csv(file = full_path, header = TRUE, sep = ",")

# dealing with the video durations
xx = gp_data$contentDetails.duration %>% ms()

yy = gp_data$contentDetails.duration %>% hms()

xx_one = xx %>% period_to_seconds()

yy_one = yy %>% period_to_seconds()

gp_data$contentDetails.duration[which(!is.na(xx))] = xx_one[which(!is.na(xx))]

gp_data$contentDetails.duration[which(!is.na(yy))] = yy_one[which(!is.na(yy))]

gp_data$contentDetails.duration = str_extract(gp_data$contentDetails.duration, "\\d+")

gp_data$contentDetails.duration = as.numeric(gp_data$contentDetails.duration)

gp_final = gp_data[,c(5,7,36,42,43,45)]

gp_final$snippet.publishedAt = gp_final$snippet.publishedAt %>% as_date

# sort from the least recent to most recent
gp_order = gp_final %>% arrange(desc(ymd(gp_final$snippet.publishedAt)))

# aggregating video statistics
sum(gp_order$statistics.viewCount)
sum(gp_order$statistics.likeCount)
sum(!is.na(gp_order$statistics.commentCount))

# dealing with videos where the comments are disabled
gp_order$statistics.commentCount[is.na(gp_order$statistics.commentCount)] = 0

# appending engagement metric into the dataset
gp_order$engagement = ( (gp_order$statistics.likeCount +
                          gp_order$statistics.commentCount)/gp_order$statistics.viewCount )*100
```


# Affiliation effect

## Exploring the views over time (between 2021 & 2022)

```{r warning=FALSE,message=FALSE}
# Libraries
library(ggplot2)
library(dplyr)
library(hrbrthemes)
library(viridis)
library(ggrepel)
library(ggthemes)
library(moments)

# Isolating 2021 and 2022
## Line plot
recent_videos = gp_order %>% 
  filter(substr(snippet.publishedAt,1,4)=="2021" | 
           substr(snippet.publishedAt,1,4)=="2022")

views_twt = recent_videos %>% ggplot(aes(snippet.publishedAt, 
                                           statistics.viewCount))

l_twt =  views_twt + geom_line(col = "black", size =1.4)+geom_smooth(method='loess',formula = 'y~x',
                  aes(color = "Trend Line"))+
  labs( x = "Date", y = "Views (Count)")+
  scale_colour_manual(name = "Legend", values = "red")+
  geom_vline(xintercept = as.numeric(recent_videos$snippet.publishedAt[233]), 
             color = "purple", linetype=2, size = 1.6) +
  annotate("text", x = recent_videos$snippet.publishedAt[229],
           y = 50000,label = "Hustler's Corner", angle = 360, size = 4) +
  geom_vline(xintercept = as.numeric(recent_videos$snippet.publishedAt[64]), 
             color = "pink", linetype=2, size = 1.6) +
  annotate("text", x = recent_videos$snippet.publishedAt[64],
           y = 50000, label = "MacG", angle = 360, size = 4) +
  ggtitle("Views plotted against time (2021 & 2022)")+
  theme_fivethirtyeight(base_size = 11, base_family = "sans")+
  theme(axis.title = element_text())

l_twt
```


## Segmenting the views data by the relevant dates
```{r message=FALSE,warning=FALSE}
#### In-depth investigation into the affiliation effect
b4_hc = gp_order %>% filter(snippet.publishedAt <= "2022-02-07") 

af_hc_b4_mcg = gp_order %>% filter(snippet.publishedAt > "2022-02-07" &
                                     snippet.publishedAt <= "2022-08-11")

af_mcg = gp_order %>% filter(snippet.publishedAt > "2022-08-11")

### Pairwise differences
pair_types = data.frame(
  groups = c(rep(1, 276), rep(2, 170), rep(3, 62)),
  views = c(b4_hc$statistics.viewCount, 
            af_hc_b4_mcg$statistics.viewCount, 
            af_mcg$statistics.viewCount)
)

pair_types
```


## Quick summary statistics

```{r warning=FALSE,message=FALSE}
library(dplyr)
pair_types %>% select(groups, views) %>%
  group_by(groups) %>%
  summarise(
    count = n(),
    mean = mean(views, na.rm = TRUE),
    sd = sd(views, na.rm = TRUE)
  )
```


## Boxplot of the different subsets of data

```{r warning=FALSE,message=FALSE}
# Plot weight by group and color by group
library("ggpubr")
ggboxplot(pair_types, x = "groups", y = "views", 
          color = "groups",
          order = c(1, 2, 3),
          ylab = "Views", xlab = "Groups")

```

## Paired t-tests

```{r warning=FALSE,message=FALSE}
with(pair_types, pairwise.t.test(views, groups))

```
Interpretations of pairwise-sample t-tests:

$$
H_{0}: \mu_{2} = \mu_{1}
\\
H_{1}: \mu_{2} > \mu_{1}
$$
The average number of views that TBP channel recieved after the 7th of Feb(the date he first appeared on the Hustler's corner with DJ Sbu) is significantly greater than before his appearance on the Hustler's corner; with probability < 0.001% that this is due to random chance.

$$
H_{0}: \mu_{3} = \mu_{2}
\\
H_{1}: \mu_{3} > \mu_{2}
$$
The average number of views that TBP channel recieved after the 11th of Aug(the date he first appeared on Podcast & Chill with MacG & Sol) is significantly greater than before his appearance on Podcast & Chill; with probability < 0.001% that this is due to random chance.


# Exploratory Data Analysis (With a focus on correlation)

## Views vs. Likes

```{r message=FALSE,warning=FALSE}
### views vs. Likes
scatter_gg = gp_order %>% ggplot(aes(x = statistics.likeCount, 
                                             y = statistics.viewCount, 
                                             color = engagement))

s_graph =  scatter_gg + geom_point()+
  labs( x = "Likes", y = "Views")+
  ggtitle("Views vs. Likes")+
  theme_fivethirtyeight(base_size = 11, base_family = "sans")+
  theme(axis.title = element_text())

s_graph + geom_smooth(method = "loess", color = "red")

```

## Views vs. Comment count

```{r message=FALSE, warning=FALSE}
scatter_two = gp_order %>% ggplot(aes(x = statistics.commentCount, 
                                              y = statistics.viewCount, 
                                              color = engagement))

s_graph_two =  scatter_two + geom_point()+
  labs( x = "Comments (Count)", y = "Views")+
  ggtitle("Views vs. Comments")+
  theme_fivethirtyeight(base_size = 11, base_family = "sans")+
  theme(axis.title = element_text())

s_graph_two + geom_smooth(method = "loess", color = "red")

```

## Views vs. Video duration

```{r message=FALSE,warning=FALSE}
scatter_three = gp_order %>% ggplot(aes(x = contentDetails.duration, 
                                      y = statistics.viewCount, 
                                      color = engagement))

s_graph_three =  scatter_three + geom_point()+
  labs( x = "Video Duration (In seconds)", y = "Views")+
  ggtitle("Views vs. Video Duration")+
  theme_fivethirtyeight(base_size = 11, base_family = "sans")+
  theme(axis.title = element_text())

s_graph_three + geom_smooth(method = "loess", color = "red")
```



# Top performing videos




<table>

Video title                                                                           Date of Upload        Views
----------                                                                         ------------------       --------
1. Rob Hersov: "Why the ANC must be removed from power & the future of South Africa      2022-08-12            65987
2. My Interview on Podcast & Chill with Mac G & Sol | Calling for Sponsors               2022-08-12            53080
3. Catching up with my old man, Joshua Maponga... speaking Dzimbabgwe & Solutions        2022-06-24            51087
4. Gayton McKenzie: A Brief Chat with The Hardest-Working Politician in SA               2022-08-15            49074
5. Vusi Thembekwayo vs Sizwe Dhlomo 🔥🔥🔥                                               2017-06-17            45776
6. Jacob Zuma was a Sensational Educator | Africans Helping South Africans | Rights      2022-09-19            38331
7. My Thoughts on The John Steenhuizen Interview on Podcast & Chill | Pol. Homework      2022-08-21            37944
8. Elias Monage of The BBC tells Ramaphosa & The ANC to go home for failing              2022-07-01            30975
9. I disagree with Vusi Thembekwayo                                                      2018-09-16            27743
10. Big Zulu & 150 Bars | Moneao | Getting Scammed by Your Partner | Reusable Pads       2022-08-22            27513














<\table>